Sketching for Density Estimation

### 2018-11-11 density.tex
row subset selection

We can also use the method of expectation minimization to achieve this bound deterministically.
Specifically, we will guaranty that $\| \sum_{j=1}^i s_j z_j \| \le \sum_{j=1}^i \|z_j\|^2$ for all $i$.
This is trivially true for $i=1$. 
For another value $i$ set $s_i = -\operatorname{sign} (\sum_{j=1}^{i-1}s_j \langle z_j, z_i \rangle)$ 
and  assume by induction that $\| \sum_{j=1}^{i-1} s_j z_j\|^2 \le \sum_{j=1}^{i-1} \|z_j\|^2$.

@@

Here are some examples:
\begin{itemize}
\item in approximate counting we have $f(x, q) = 1$ if $x=q$ and zero else.
\item in quantile approximation we have $f(x, q) = 1$ if $x<q$ and zero else.
\item in matrix column subset selection we have $f(x, q) = \langle x,q \rangle ^2$.
\item in density estimation $f(x, q) = \exp(- \|x-y\|_2^2/\sigma^2)$ or any other kernel (see above)
\item in machine learning $f(x, q) = \operatorname{loss}(x,q)$ where $q$ is the parameters of the model and $x$ is a training example and its label.
\end{itemize}
For all of the above, we can find flat error functions $E(q) = \sum_{i=1}^{k} s_i f(x_i,q)$ such that $\sup_q E(q)$ is small. 
We argue that this gives a general mechanism for sketching in a diverse set of scenarios. 
\end{document}

Let us choose the value  and assume by induction that

>>>>

We can also use the method of expectation minimization to achieve this bound deterministically.
Specifically, we will guaranty that $\| \sum_{j=1}^i s_j z_j \|^2 \le \sum_{j=1}^i \|z_j\|^2$ for all $i$.
This is trivially true for $i=1$. 
For another value $i$ set $s_i = -\operatorname{sign} (\sum_{j=1}^{i-1}s_j \langle z_j, z_i \rangle)$ 
and  assume by induction that $\| \sum_{j=1}^{i-1} s_j z_j\|^2 \le \sum_{j=1}^{i-1} \|z_j\|^2$.

@@

Here are some examples:
\begin{itemize}
\item in approximate counting we have $f(x, q) = 1$ if $x=q$ and zero else.
\item in quantile approximation we have $f(x, q) = 1$ if $x<q$ and zero else.
\item in matrix row subset selection we have $f(x, q) = \langle x,q \rangle ^2$.
\item in density estimation $f(x, q) = \exp(- \|x-y\|_2^2/\sigma^2)$ or any other kernel (see above)
\item in machine learning $f(x, q) = \operatorname{loss}(x,q)$ where $q$ is the parameters of the model and $x$ is a training example and its label.
\end{itemize}
For all of the above, we can find flat error functions $E(q) = \sum_{i=1}^{k} s_i f(x_i,q)$ such that $\sup_q E(q)$ is small. 
We argue that this gives a general mechanism for sketching in a diverse set of scenarios.

Let us choose the value  and assume by induction that
\section{Row subset selection}
Assume you are getting the rows $x_i$  of matrix $X$ in a stream and you want to compute $Z$ such that $\|Z^TZ - X^TX\| \le \eps n$.
for simplicity, assume $x_i$ are all unit length. Moreover, assume each vector $Z$ must be one of the rows in $X$ up to some constant factor.
This is called the row subset selection problem. 
Since you are trying to minimize the quadratic form, you get that $f(x, q) = \langle x,q \rangle ^2$ and 
$$E(y) = \sum_i s_i \langle x_i,q \rangle ^2 = q^T (\sum_i s_i x_i x_i^T ) q \le \|\sum_i s_i x_i x_i^T\|$$
It is clear that we can get $\|\sum_i s_i x_i x_i^T\| \le \sqrt{k \log(d)}$. Could it would see that maybe $O(1)$ is also possible?
When summing rank one projections matrices $x_i x_i^T$ you can think of two extreme cases. 
If $x_i$ are orthogonal to each other than $\|\sum_i s_i x_i x_i^T\| = 1$ independently of $s_i$.
If $x_i$ are all equal than a random assignment would get  $\E[\|\sum_i s_i x_i x_i^T\|] = E[|\sum_i s_i|] = \Omega(\sqrt{k})$.
But, in that case it's also easy to find a sequence of $s_i$ which gives $\|\sum_i s_i x_i x_i^T\| = O(1)$. Is that always possible?
If so, it would give a covariance sketch is subset selection of size $1/\eps$ which would be an improvement over frequent directions.

\end{document}
